Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-12 doc/report/msc-report.tex
doc/report/msc-report.tex: small notes for metadata.
doc/report/msc-report.tex: intermediate push for introduction.

\subsection{Scope}

@@

@@

\chapter{Theory}

\chapterquote{Einstein explained his theory to me every day, and on my
arrival I was fully convinced that he understood it.}{Chaim Weizmann}

\section{Other data warehouses}

@@

\chapter{Tools}

%\chapterquote{In short, intelligence, considered in what seems to be its
%original feature, is the faculty of manufacturing artificial objects,
%especially tools to make tools, and of indefinitely urging the
%manufacture.}{Henri Bergson}

@@

% TODO> I argue that we should have 'hadoop' and 'hbase' as the main sections,
% and let mapreduce be a subsection of hadooop, bigtable a dito of hbase.

\section{MapReduce}

@@

Basically MapReduce is a two phase computation. First the input data is
partitioned and each of these partitions is sent to a \textit{mapper} which
performs computations on each row of the data. The mapper then outputs each
resulting computation on the format \texttt{key,value}. These key-value
pairs are received by a \textit{reducer} which performs the second stage of
the computation.

\section{BigTable}

@@

\subsubsection{Rows}

@@

\subsubsection{Regions}

@@

% TODO> why is this Tool & BigTable-subsection called Dimensions?
% Disambiguation needed. -fredrik
\subsection{Dimensions}

\subsubsection{Column-families}

@@

\subsubsection{Column-qualifiers}

@@

\subsection{Hierarchies}

@@

The minimum cost of scanning data is achieved by ordering the data so that
a scan uses every visited row. In the case with thousand rows of which ten
are wanted, would mean that ten rows scan are scanned of which all are
used.

All Zohmg import programs are executed on Hadoop.

\subsection{Thrift}

@@

\section{Dumbo}

Dumbo is a framework for writing Hadoop programs in languages other than
Java, in a so-called Streaming Mode. Dumbo is written in Python. It is used
extensively at Last.fm for writing short prototypes. \cite{dumbo}

\section{Serializing}

@@

\subsection{Paste}

@@

@@

%\chapterquote{It is not always what we know or analyzed before we make a
%decision that makes it a great decision. It is what we do after we make the
%decision to implement and execute it that makes it a good
%decision.}{William Pollard}

@@

 yield log.timestamp, dimensions, measurements
\end{verbatim}

@@

@@

@@

This is the user manual.

A manage script is used to create tables from a pre-defined schema.

\noindent Edit the configuration files, which are text files in YAML and
Python format.

\noindent The above command executes the configuration made and creates the
infrastructure (HBase tables) for the project.

The user writes the map part of the MapReduce programs which imports
log-like data to the data store. The reduce step is taken care of by Zohmg,
which uses sum reducer that simply sums up values with the same keys.

\subsection{User mappers}

Import jobs use mappers written by the user, either in Python or Java.
Mappers written by the user import each line from the input file and set
the arguments to the map function, key and value, as line number and line
contents, respectively. The user can then perform computations on these
entities and finally yield the results as a dictionary. The resulting
dictionary will be interpreted by Zohmg and stored accordingly.

\section{Storing data}

% COMMENT: I don't think we should mention the underlying store in the usage section.
Data is stored in Apache HBase. A natural consequence of this that the
data is stored sorted by rowkey allowing fast access to an arbitrary
(sequential) range.

Data is served to the user either over HTTP in JSON format or over Thrift.

\noindent The above serves on default port 8086, with the \texttt{--port}
argument it is possible to change server port.

\subsection{Data}

Based on the query---which the user submits---the corresponding data is
served from HBase as JSONP. The query is a simple definition of wanted row
range, dimensions, and values for these dimensions. The query is submitted
to the data server via HTTP, an example query could look like (the URL is
wrapped to fit the page)

The above query would return the pageviews for SE between 1 January 2009
and 4 April 2009.

\subsection{Transformers}

\subsection{Clients}

The data server serves static files from the project's \texttt{clients}
directory at the server URL appended with \texttt{/client/filename}.

\chapter{Examples}

\chapterquote{Example is not the main thing in influencing others, it's
the only thing.}{Albert Schweitzer}

%\chapterquote{There is nothing so annoying as a good example!!}{Mark Twain}

%\chapterquote{Example isn't another way to teach, it is the only way to
%teach}{Albert Einstein}

\section{Web log analysis}

Configure the dataset.yaml.

Write a custom mapper, which imports data from web logs.

Start data server and extract data from it.

Possibly write a transformer which performs some transformation on the
aggregated data.

Possibly write a client which will render a graph.

\chapter{Discussion}

\chapterquote{In discussion comes the light!}{Unknown}

\section{Related work}

@@

It'd be interesting to think a bit about how the data store could be fully
elastic, i.e. not require any data set definition from the user and instead
create column-families on the fly based on the dimensions output from the
user's map.
\subsection{Multiple data sets}

@@

@@

\chapter{Conclusion}

Last.fm is looking at using Zohmg for multiple application areas, such as
radio stats and web page stats.

@@

>>>>

Last.fm

SQL

EXPLODE

\section{Scope}

@@

\subsubsection*{Related work}

Omniture is used at Last.fm etc.

\section{Goal}

The goal (sv. syfte) with our work.

@@

\chapter{Analysis}

\section{Task}

\subsection{Import}

\subsection{Store}

\subsection{Export}

\section{Interface}

@@

\section{System}

@@

\chapter{Method}

\subsection{MapReduce}

@@

Basically MapReduce is a two phase computation. First the input data is
partitioned and each of these partitions is sent to a \textit{mapper} which
performs computations on each row of the data. The mapper then outputs each
resulting computation on the format $\langle$\textit{key,value}$\rangle$. These
key-value pairs are received by a \textit{reducer} which performs the second
stage of the computation.
All Zohmg import programs are executed on Hadoop.

\subsection{Dumbo}

Dumbo is a Python framework for writing Hadoop programs in languages other than
Java, in a so-called Streaming Mode. It is used extensively at Last.fm for
writing short prototypes. \cite{dumbo}

\subsection{BigTable}

@@

\subsubsection*{Rows}

@@

\subsubsection*{Regions}

@@

% TODO> why is this Tool & BigTable-subsection called Dimensions?
% Disambiguation needed. -fredrik
\subsection*{Dimensions}

The degrees of freedom in BigTable.

\subsubsection*{Column-families}

@@

\subsubsection*{Column-qualifiers}

@@

\subsection*{Hierarchies}

@@

The minimum cost of scanning data is achieved by ordering the data so that
a scan uses every visited row. In the case with thousand rows of which ten
are wanted, would mean that ten rows scan are scanned of which all are
used.
\subsection{Thrift}

@@

@@

@@

Data sets and environment.

\section{Exporting data}

TODO: More general about middleware, move Paste mentioning to Implementation
chapter.

@@

@@

    yield log.timestamp,dimensions,measurements
\end{verbatim}

@@

Typed bytes

@@

TextInputFormat

LzoInputFormat

@@

\chapter{Results}

\section{Discussion}

@@

It'd be interesting to think a bit about how the data store could be fully
elastic, i.e. not require any data set definition from the user and instead
create column-families on the fly based on the dimensions output from the
user's map.

\subsection{Multiple data sets}

@@

data set introspection

roll-up and drill-down

good when creating apps

@@

\section{Conclusion}

Last.fm is investigating the possibility for using Zohmg for multiple
application areas, such as radio stats and web page stats.

@@

\appendix

\noindent The following chapter will describe how the Zohmg system is used.
Project deployment and data import and export will be described briefly.

The Zohmg manage script is used to create a project directory

\noindent The next step is editing the configuration files, which are located in
the project's \texttt{config} directory. The configuration files are text files
in YAML and Python format. When the files are edited, run the setup command;
this executes the configuration and creates infrastructure (HBase tables) for
the project.

In order to import log-like data to the data store, the map part of the
MapReduce program needs to be created. The reduce step is taken care of by
Zohmg, which uses sum reducer that simply sums up values with the same keys.

\subsection*{User mappers}

Import jobs use mappers written either in Python or Java. User mappers import
each line from the input file and set the arguments to the map function, key and
value, as line number and line contents, respectively. The user can then perform
computations on these entities and finally yield the results as a dictionary.
The resulting dictionary will be interpreted by Zohmg and stored accordingly.

Data is served to the user over HTTP in JSONP format or. To start the data
server execute

The command above starts the data server on default port 8086, with the
\texttt{--port} argument it is possible to change server port.

\subsection*{Data}

Based on the query---which the user submits---the corresponding data is
served from HBase as JSONP. The query is a simple definition of wanted row
range, dimensions, projections, or both, and values for these. The query is
submitted to the data server via HTTP, a query could look like

\noindent The above query would return the pageviews for SE between 1 January
2009 and 4 April 2009.

\subsection*{Transformers}

Accessing transformed data is done similarly to the raw data extraction.

\begin{verbatim}
   http://localhost:8086/transform/transformer/
   ?t0=20090101&t1=20090420&unit=pageviews
   &d0=country&d0v=SE
\end{verbatim}

\noindent The above query extracts the same data as the query from the Data
section. This data is run through the \texttt{transformer.py} application and
is, again, served to the user as JSONP.

\subsection*{Clients}

The data server serves static files from the project's \texttt{clients} from

\begin{verbatim}
   http://localhost:8086/clients/filename
\end{verbatim}

\chapter{Glossary}

\begin{itemize}
\item \texttt{Serializing}
